\thetitle

### 2013-09-29 tfpie/response.txt

----------------------- REVIEW 1 ---------------------

I use the term "work-in-progress," because the authors have
side-stepped the methodology used to introduce types and the
transition from their tower of student languages to an industrial
language. Granted, this is likely something the authors are still
polishing. Nonetheless, it is important to provide a vision for how to
do these things and to outline what they have done highlighting both
succeses and failures. In this manner, interested readers may join
their efforts by reproducing the authors' methodology in their schools
using the languages developed by the authors. Let's face it,
introducing types is necessary to make the move to an
industrial-strength language and without providing a vision for it
many may balk. I suspect that the transition builds on the informal
contracts students are familiar with. Thus, adding this vision should
not be to difficult. Outlining this vision may obviate doubts readers
may have about making the transition to an industrial-strength
language. This is the primary reason that I am unable to put this
article in the strong accept column.

RESPONSE: We have added a paragraph in section 4.1 on this subject.

### 2013-09-29 tfpie/tfpie.tex
sync w/ david

Many universities and colleges aim to teach their students proficiency
in an industrial object-oriented programming language by the end of
the students' first year.  The most common approache to achieve this
aim is to teach an industrial language in the first course, often
Java, starting on the first day.  Other curricula take a more indirect
route by teaching functional programming in the first semester,
followed by a second semester in Java.  The latter approach is an
improvement over the first, as pointed out by numerous
observers~\cite{dvanhorn:Felleisen2004Structure, local:spolsky}, but
both suffer serious flaws.

@@

This abrupt transition has several negative consequences: the
principles of object-oriented programming are obscured and
de-emphasized, struggling with the programming environment is
frustrating and can cause potentially good students to leave the
program, it favors students with prior exposure to the particular
tools (a set that is demographically skewed), it inhibits students
from experimenting by allowing them to rely upon past skills, and it creates
the false impression that courses are discrete units of instruction
that can be discarded after successful completion rather than being
part of a continuous and cumulative educational experience.

@@

Unions, and recursive unions in particular, are a fundamental kind of
data definition that students are well-versed in from the previous
semester.  A fundamental early lesson is how to represent (recursive)
unions using classes and how to write recursive methods.  As an
example, figure~\ref{fig:tree} defines binary trees of numbers
(an archetypal recursive union data definition)
using the BSL language and the Class language.
%% \begin{verbatim}
%%     ;; A list of numbers (LoN) is one of:
%%     ;; - (new empty%)
%%     ;; - (new cons% Number LoN)

>>>>

Many universities and colleges aim to teach their students proficiency
in an industrial object-oriented programming language by the end of
the students' first year.  The most common approach to achieve this
aim is to teach an industrial language in the first course, often
Java, starting on the first day.  Other curricula take a more indirect
route by teaching functional programming in the first semester,
followed by a second semester in Java.  The latter approach is an
improvement over the first, as pointed out by numerous
observers~\cite{dvanhorn:Felleisen2004Structure, local:spolsky}, but
both suffer serious flaws.

@@

This abrupt transition has several negative consequences which we have
experienced first-hand: the principles of object-oriented programming
are obscured and de-emphasized, struggling with the programming
environment is frustrating and can cause potentially good students to
leave the program, it favors students with prior exposure to the
particular tools (a set that is demographically skewed), it inhibits
students from experimenting by allowing them to rely upon past skills,
and it creates the false impression that courses are discrete units of
instruction that can be discarded after successful completion rather
than being part of a continuous and cumulative educational experience.

@@

Unions, and recursive unions in particular, are a fundamental kind of
data definition that students are well-versed in from the previous
semester.  A fundamental early lesson is how to represent (recursive)
unions using classes and how to write recursive methods.  As an
example, figure~\ref{fig:tree} defines binary trees of numbers
(an archetypal recursive union data definition)
using the Beginning Student language (BSL) as used at the start of the
first semester, and also using the Class language of our course.
%% \begin{verbatim}
%%     ;; A list of numbers (LoN) is one of:
%%     ;; - (new empty%)
%%     ;; - (new cons% Number LoN)
